---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# getsat: Retrieving Commonly Used Satellite Data

<!-- badges: start -->
<!-- badges: end -->

`getsat` is an R package designed to simplify access to commonly used satellite data on environmental variables. It provides an easy way to retrieve, process, and integrate satellite-based data from sources such as:
- Copernicus Atmosphere Monitoring Service
- Copernicus Climate Data Store
- Microsoft Planetary Computer


## Installation

To install the development version of `getsat` from [GitHub](https://github.com/), use the [pak](https://CRAN.R-project.org/package=pak) package:
```r
# Install pak if not already installed
install.packages("pak")

# Install getsat from GitHub
pak::pak("jalilian/getsat")
```

## Example 1: Retrieve Elevation Data

The following example shows  how to retrieve elevation data from the Copernicus Digital Elevation Model (DEM). You can specify a bounding box for an area of interest or provide specific coordinates.
```{r example1}
library("getsat")

# Retrieve elevation data for a bounding box (longitude/latitude)
elev1 <- get_dem(c(47, 34, 47.5, 35), res=90)
# plot retrived elevation data
terra::plot(elev1)

# Retrieve elevation for specific coordinate points
coords <- cbind(runif(n = 100, min = 47, max = 47.5), 
                runif(n = 100, min = 34, max = 35))
elev2 <- get_dem(coords, res = 90)
# Print retrieved elevation values
head(elev2)
```

## Example 2: Retrieve Elevation Data

This example shows how to retrieve 8-day daytime temperature data from the Moderate Resolution Imaging Spectroradiometer (MODIS). You can specify a bounding box or coordinates, and provide the desired time period in the format "YYYY-MM-DD/YYYY-MM-DD".
```{r example2}
# Retrieve temperature data for a bounding box (longitude/latitude) and specific time period
temp <- get_modis(c(47, 34, 47.5, 35), var = "LST_Day_1KM",
                  datetime = "2024-12-01/2025-01-28")
# plot retrived temperature data
terra::plot(temp)
```
