---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# getsat: Retrieving Commonly Used Satellite Data

<!-- badges: start -->
[![R Build Status](https://github.com/jalilian/getsat/workflows/R-CMD-check/badge.svg)](https://github.com/jalilian/getsat/actions)
<!-- badges: end -->

`getsat` is an R package designed to simplify access to commonly used satellite data on environmental variables. It provides an easy way to retrieve, process, and integrate satellite-based data from sources such as:

- [Copernicus Climate Data Store](https://cds.climate.copernicus.eu)
- [Copernicus Atmosphere Data Store](https://ads.atmosphere.copernicus.eu)
- [Microsoft Planetary Computer](https://planetarycomputer.microsoft.com)
- [Climatology Lab](https://www.climatologylab.org/terraclimate.html)

Please note that users must acknowledge the original data sources appropriately. Any use of data from these platforms requires clear and visible attribution, including proper citation and referencing of the datasets, as specified by their licenses and terms of use.

Development of this package was supported by the Lancaster Ecology and Epidemiology Group (LEEG) and funded by Biotechnology and Biological Sciences Research Council (BB/Y514238/1) and the Wellcome Trust (220870/Z/20/Z).


## Installation

To install the development version of `getsat` from [GitHub](https://github.com/), use the [remotes](https://CRAN.R-project.org/package=remotes) package:
```r
# Install pak if not already installed
install.packages("remotes")

# Install getsat from GitHub
remotes::install_github("jalilian/getsat")
```

## Example 1: Retrieve elevation data

The following example shows  how to retrieve elevation data from the [Copernicus Digital Elevation Model (DEM)](https://doi.org/10.5069/G9028PQB). You can specify a bounding box for an area of interest or provide specific coordinates.
```{r example1}
library("getsat")

# Retrieve elevation data for a bounding box (longitude/latitude)
elev1 <- get_dem(c(47, 34, 47.5, 35), res=90)
# plot retrived elevation data
terra::plot(elev1)

# Retrieve elevation for specific coordinate points
coords <- cbind(runif(n = 100, min = 47, max = 47.5), 
                runif(n = 100, min = 34, max = 35))
elev2 <- get_dem(coords, res = 90)
# Print retrieved elevation values
head(elev2)
```


## Example 2: Retrieve temprature from MODIS

This example shows how to retrieve 8-day daytime temperature data from the [Moderate Resolution Imaging Spectroradiometer (MODIS)](https://planetarycomputer.microsoft.com/dataset/group/modis). Make sure to properly cite MODIS data following [NASA's guidelines](https://modaps.modaps.eosdis.nasa.gov/services/faq/LAADS_Data-Use_Citation_Policies.pdf). You can specify a bounding box "c(xmin, ymin, xmax, ymax)" or  coordinates (longitude, latitude), and provide the desired time period in the format "YYYY-MM-DD/YYYY-MM-DD".
```{r example2}
# Retrieve temperature data for a bounding box (longitude/latitude) and specific time period
temp <- get_modis(c(47, 34, 47.5, 35), var = "LST_Day_1KM",
                  datetime = "2025-01-01/2025-03-09")
# plot retrived temperature data
terra::plot(temp)
```

